Is There a Competitive Advantage to Using Multivariate Statistical or Machine Learning Methods Over the Bross Formula in the hdPS Framework for Bias and Variance Estimation?

### 2024-09-14 Writeup/manuscript.Rmd
bmc submission

# Sample for structured abstract
abstract: |
  **Purpose**: We aims to evaluate various proxy selection methods within the context of high-dimensional propensity score (hdPS) analysis. The study focuses on assessing the performance of these methods, including alternative variable selection approaches, in selecting proxy variables for confounding adjustment compared to the traditional hdPS method that is rooted in Bross formula. The goal is to understand better the performance of these alternative methods in estimating treatment effects, and identify scenarios in which they may perform better.
  **Methods:** Using data from three cycles of the National Health and Nutrition Examination Survey (NHANES) spanning 2013-2018, we motivated the study by examining the association between obesity and diabetes. A plasmode simulation framework based on this data was employed to mimic real-world data structures. Simulations were conducted under three scenarios: Frequent Exposure and Outcome, Rare Exposure and Frequent Outcome, and Frequent Exposure and Rare Outcome. The performance of a variety of proxy selection methods—including tree-based methods, LASSO-based methods, and the Genetic Algorithm (GA)—was evaluated across these scenarios using standard simulation metrics.
  **Results:** XGBoost consistently demonstrated the lowest MSE and high coverage, making it a reliable method overall, although it did not always exhibit the lowest bias. In contrast, GA consistently showed the highest bias and MSE, with lower coverage and greater variability, indicating its unsuitability for accurate effect estimation. The kitchen sink model, Bross-based hdPS, and Hybrid hdPS methods performed moderately well, with low bias and moderate MSE, but varied in coverage across scenarios. Scenario-specific trends revealed that rare outcome scenarios yielded lower MSE and better precision, while rare exposure scenarios were associated with higher bias and MSE.
  **Conclusion:** Aside from GA, the performance of most other methods in your study appears to be comparable in each scenario, though with some variations depending on specific metrics. The findings underscore the importance of selecting appropriate methods for hdPS analysis based on the characteristics of the data, particularly the prevalence of exposure and outcome. While XGBoost excelled in overall accuracy and precision, the choice of method should be tailored to the specific epidemiological goal to optimize bias, coverage, and precision in estimating treatment effects.

@@

**High-dimensional Propensity Score (hdPS) Algorithm**: In epidemiology, proxy variables are commonly used as substitutes for confounders that are difficult or impossible to measure directly, such as socioeconomic status, lifestyle factors, or health behaviors [@vanderweele2019principles]. The high-dimensional propensity score (hdPS) is a pharmacoepidemiological method designed to reduce confounding bias in large healthcare databases [@schneeweiss2009high]. Unlike traditional propensity score models that rely on investigator-specified or manually selected covariates, hdPS automatically ranks a wide array of proxy variables from healthcare records—such as diagnosis codes, medications, and procedures—using the Bross formula [@wyss2018erratum; @bross1966spurious]. The Bross formula ranks these variables based on their marginal associations with both exposure and outcome. These selected proxy variables serve as surrogates for unmeasured or poorly measured confounders, helping reduce bias in treatment effect estimates. The hdPS algorithm further refines the selection by prioritizing variables based on their prevalence and potential for confounding, defined by their association with both exposure and outcome [@schneeweiss2009high].

**Multivariate Machine Learning Extensions**: Although the Bross formula performs well in certain contexts, it has limitations in capturing complex interactions, nonlinearities, and higher-order associations between variables, especially in high-dimensional settings where it does not account for the multivariate structure of other covariates [@karim2024high; @karim2018can]. To address these limitations, multivariate machine learning methods such as LASSO, Elastic Net, and Random Forests have been applied within the hdPS framework. These methods are better suited for high-dimensional data, where they can more effectively handle complex relationships and improve the selection of proxy variables, thus enhancing the precision of treatment effect estimates [@karim2018can; @schneeweiss2017variable; @franklin2015regularized]. Simulation studies and empirical research have shown that these machine learning-based methods, or hybrid approaches combining the Bross formula with machine learning, can reduce confounding more effectively and increase efficiency compared to the Bross formula alone in certain settings [@karim2018can; @schneeweiss2017variable; @franklin2015regularized].

**Assessing the Simulation Performance**: Other than bias, previous studies have primarily focused on Mean Squared Error (MSE) as a key metric for evaluating the performance of hdPS and its machine learning extensions [@franklin2015regularized; @pang2016effect; @wyss2018using; @karim2018can; @simon2023evaluating]. However, in high-dimensional settings with singly robust methods, such as hdPS and machine learning approaches like LASSO, MSE may not always be the most reliable measure. MSE combines both bias and variance into a single metric, which makes interpretation challenging when variance estimation is unstable—a common issue in these methods. In contrast, coverage, which measures the proportion of confidence intervals that capture the true treatment effect, provides a more direct and meaningful assessment of a model's reliability. In realistic observational studies, where model misspecification is often inevitable, coverage—along with related metrics such as bias-eliminated coverage and relative error in standard error (SE; which compares model-based SE with empirical SE)—can reveal whether confidence intervals or SEs are too narrow (underestimating uncertainty) or too wide (overestimating uncertainty). This insight is crucial in determining whether the model delivers valid estimates despite misspecification. Even a model with poor MSE but good coverage may still be valuable, as it produces realistic confidence intervals. By shifting the focus to coverage, rather than relying solely on MSE, we can achieve a more comprehensive understanding of method performance, especially in cases where unstable variance estimation might distort conclusions drawn from MSE alone.

@@

**Plasmode simulation**: To rigorously assess the performance of the methods under consideration, we employed a plasmode simulation framework, which is particularly well-suited for reflecting real-world data structures and complexities [@franklin2014plasmode]. This approach was modeled after the analytic dataset derived from NHANES and involved resampling from the observed covariates and exposure information (i.e., obesity) without altering them. By mirroring key aspects of an actual epidemiological study, this simulation framework offers a substantial advantage over traditional Monte Carlo simulations, which often rely on hypothetical assumptions.

@@

**Performance Measures**: From this simulation, we derived several performance metrics to evaluate the effectiveness of the methods under consideration: (1) bias, (2) average model standard error (SE; the average of estimated SEs obtained from a model over repeated samples), (3) empirical SE (the standard deviation of estimated treatment effects across repeated samples), (4) mean squared error (MSE), (5) coverage probability of 95% confidence intervals, (6) bias-corrected coverage, and (7) Zip plot [@morris2019using; @white2023check].

@@

4. **Hybrid of hdPS and LASSO**: Instead of relying solely on LASSO for variable selection, a hybrid approach combines the Bross formula and LASSO. First, hdPS variables are selected using the hdPS algorithm (e.g., the top 100), and then LASSO is applied to further refine the selection [@karim2018can; @franklin2015regularized].

@@

6. **Random Forest**: The Random Forest (RF) algorithm is an ensemble learning method that constructs multiple decision trees to perform classification [@breiman2001random]. It calculates the importance of each proxy variable based on the decrease in impurity or Gini importance, providing a ranking of the proxies. The top 100 variables from this ranking are manually selected for further modeling [@schneeweiss2017variable].

@@

9. **Genetic algorithm (GA)**: GA is an evolutionary algorithm inspired by the theory of natural selection [@holland1975adaptation]. It operates by evolving offspring from a population of the fittest individuals over several generations, evaluating and selecting the best combination of features or variables that maximize prediction accuracy.

@@

The results for each method under the different scenarios are summarized below. See Figures \ref{fig:bias-comparison} and \ref{fig:Coverage-comparison} for an overview of the performance in terms of bias and coverage, respectively.

1. *Bias*: Bross-based hdPS exhibited the smallest bias (-0.0001), and the kitchen sink model (0.0002) was the second. GA shows the highest bias (0.0287), indicating a substantial deviation from the true effect. Among the other methods, Hybrid hdPS (0.0016), and Elastic Net (0.0036) demonstrated low bias. XGBoost (0.0074) had slightly higher bias than Random Forest (RF) (0.0034).

2. *Coverage*: The coverage for most methods was high, with Hybrid hdPS, Forward Selection, Backward Elimination, LASSO, and Elastic Net achieving values around 98%, indicating well-calibrated confidence intervals. However, GA had noticeably lower coverage (83.8%), indicating that its confidence intervals might be too narrow or biased, potentially missing the true effect. After applying bias elimination (as there were substantial bias associated with this method), GA's coverage improved to 96%.

3. *Mean Squared Error (MSE)*: XGBoost achieved the lowest MSE (0.0006). GA maintained the highest MSE (0.0016), reflecting its higher bias and variability. The kitchen sink model (0.0009), Bross-based hdPS (0.0008), Hybrid hdPS (0.0008), and Elastic Net (0.0009) all had relatively similar and moderate MSE values.

4. *Standard Error (SE)*: XGBoost exhibited the lowest Empirical SE (0.0229), indicating high precision in its estimates. The kitchen sink model had the highest Empirical SE (0.0305), suggesting greater variability. Other methods, including GA (0.0274), Hybrid hdPS (0.0278), and Bross-based hdPS (0.0287), showed moderate variability. LASSO (0.0299) and Elastic Net (0.0294) had slightly higher variability. The Model-based SE followed a similar pattern, with XGBoost (0.0268) showing the lowest variability and the kitchen sink model (0.0333) the highest, indicating less precision in its estimates. When comparing relative error in SE estimation, XGBoost performed the worst. See Appendix \S C for further details.

1. *Bias*: The kitchen sink model showed a relatively low bias (0.0025), while GA continued to exhibit the highest bias (0.0408), indicating a significant deviation from the true effect. XGBoost had a bias of 0.0259, which, while still higher than other methods, but was lower than GA. Other methods, such as Bross-based hdPS (0.0035), Hybrid hdPS (0.0049), and Elastic Net (0.0053), demonstrated moderate bias. Random Forest (RF) (0.0127) and Forward Selection (0.0108) had slightly higher bias but remained within an acceptable range.

2. *Coverage*: Coverage levels remained high for most methods, with XGBoost achieving the highest coverage at 96.2%, indicating well-calibrated confidence intervals despite its higher bias. The GA method had lower coverage (92.2%), suggesting that its confidence intervals might be narrower, potentially missing the true effect. Other methods such as RF, Forward Selection, Backward Elimination, and Hybrid hdPS maintained coverage values around 94-95%, suggesting adequate interval calibration. Bias-eliminated coverage for GA improved to 94.2%, but it was still slightly lower than other methods (e.g., forward selection).

3. *Mean Squared Error (MSE)*: Forward selection demonstrated the lowest MSE (0.0030), and then Hybrid hdPS and XGBoost. The GA method had a higher MSE (0.0043), reflecting its substantial bias and variability. The kitchen sink model also had an MSE of 0.0043, similar to GA, while other methods like Bross-based hdPS (0.0035), RF (0.0039), and Elastic Net (0.0036) exhibited moderate MSE values, indicating reasonable accuracy.

4. *Standard Error (SE)*: The lowest Empirical SE was observed with XGBoost (0.0507) and GA (0.0510), reflecting high precision despite their higher bias. The kitchen sink model had the highest Empirical SE (0.0656), indicating greater variability. Hybrid hdPS (0.0564), Bross-based hdPS (0.0595), and RF (0.0609) showed moderate variability. Forward Selection (0.0537) and Backward Elimination (0.0576) had lower variability compared to the kitchen sink model. In terms of Model-based SE, XGBoost (0.0531) and GA (0.0533) continued to show low variability, while the kitchen sink model had the highest Model-based SE (0.0623), indicating less precision in its estimates. When comparing relative error in SE estimation, XGBoost and kitchen sink model performed the worst (in other direction).

1. *Bias*: In this scenario, the kitchen sink model exhibited a moderate negative bias (-0.0093), similar to the Bross-based hdPS method (-0.0088). GA showed a significantly higher bias (0.0362), indicating a substantial deviation from the true effect. Among other methods, XGBoost demonstrated the lowest bias (-0.0061), while methods like Hybrid hdPS (-0.0082), Forward Selection (-0.0070), and Backward Elimination (-0.0070) had slightly higher but still moderate biases. Elastic Net and LASSO both had biases of -0.0079, reflecting slightly larger deviations compared to XGBoost but still within acceptable limits.

2. *Coverage*: Most methods achieved good coverage, with XGBoost, RF, and Forward Selection each achieving a coverage rate of 95.4%, indicating well-calibrated confidence intervals. The GA method, however, had slightly lower coverage (91.8%), indicating that its confidence intervals might be narrower, potentially excluding the true effect. Bross-based hdPS and the kitchen sink model had slightly lower coverage values of 93.8% and 93.4%, respectively. After accounting for bias, the bias-eliminated coverage for most methods, except GA, remained high, with values ranging from 98.4% to 99.0%, indicating that most methods effectively adjusted for bias in their coverage estimates. GA's bias-eliminated coverage was lower at 93.4%, reflecting its higher inherent bias.

3. *Mean Squared Error (MSE)*: XGBoost exhibited the lowest MSE (0.0003). GA had the highest MSE (0.0040), reflecting its substantial bias and variability. The kitchen sink model (0.0005), Bross-based hdPS (0.0005), and other methods like Hybrid hdPS (0.0004) and Elastic Net (0.0005) all had relatively similar MSE values, indicating moderate accuracy.

4. *Standard Error (SE)*: The lowest Empirical SE was observed with XGBoost (0.0152), reflecting high precision in its estimates. The GA method exhibited the highest Empirical SE (0.0523), indicating greater variability and less precision. Methods such as Hybrid hdPS (0.0184), Forward Selection (0.0187), and Elastic Net (0.0203) showed moderate variability, while Bross-based hdPS (0.0206) and the kitchen sink model (0.0212) had slightly higher variability. In terms of Model-based SE, XGBoost (0.0179) again showed the lowest variability, consistent with its low Empirical SE, indicating that it provided the most stable estimates. The kitchen sink model had a slightly higher Model-based SE (0.0219), indicating less precision in its estimates. When comparing relative error in SE estimation, XGBoost performed the worst.

@@

@@

# Combine RD and OR data into a single data frame
data_combined <- data.frame(
  Method = rep(c("KitchenSink", "Bross", "Hybrid", "LASSO", "ElasticNet", "RF", "XGBoost", "FS", "BE", "GA"), 2),
  # These are approximate values
  Value = round(c(0.075, 0.077, 0.079, 0.078, 0.075, 0.085, 0.082, 0.078, 0.078, 0.076, 
            1.502, 1.522, 1.543, 1.531, 1.504, 1.588, 1.564, 1.533, 1.533, 1.515),2),
  LowerCI = round(c(0.056, 0.058, 0.06, 0.059, 0.056, 0.066, 0.063, 0.059, 0.059, 0.057,
              1.395, 1.413, 1.432, 1.421, 1.396, 1.474, 1.452, 1.423, 1.423, 1.406),2),
  UpperCI = round(c(0.095, 0.096, 0.099, 0.098, 0.094, 0.104, 0.101, 0.097, 0.097, 0.096,
              1.618, 1.64, 1.663, 1.65, 1.62, 1.712, 1.685, 1.653, 1.653, 1.632),2),
  Measure = rep(c("RD", "OR"), each = 10),
  Variables = rep(c(142, 100, 49, 60, 69, 100, 48, 59, 59, 64), 2)
)

@@

\begin{table}[htbp]
\centering
\caption{Comparison of variable overlap of selected proxies across different methods used to evaluate the association between obesity and diabetes}
\label{tab:method-comparison-updated}
\begin{tabular}{lcccccccccc}
\toprule
 & \textbf{KS} & \textbf{Bross} & \textbf{Hybrid} & \textbf{LASSO} & \textbf{EN} & \textbf{RF} & \textbf{XGB} & \textbf{FS} & \textbf{BE} & \textbf{GA} \\
\midrule
\textbf{Kitchen sink (KS)} & 142 & & & & & & & & & \\
\textbf{Bross formula} & 100 & 100 & & & & & & & & \\
\textbf{Hybrid (Bross and LASSO)} & 49 & 49 & 49 & & & & & & & \\
\textbf{LASSO} & 60 & 47 & 47 & 60 & & & & & & \\
\textbf{Elastic Net (EN)} & 69 & 54 & 48 & 60 & 69 & & & & & \\
\textbf{Random Forest (RF)} & 100 & 71 & 38 & 46 & 52 & 100 & & & & \\  
\textbf{XGBoost (XGB)} & 48 & 38 & 24 & 28 & 30 & 37 & 48 & & & \\
\textbf{Forward selection (FS)} & 59 & 45 & 41 & 51 & 54 & 45 & 25 & 59 & & \\
\textbf{Backward elimination (BE)} & 59 & 45 & 41 & 51 & 54 & 45 & 25 & 59 & 59 & \\
\textbf{Genetic algorithm (GA)} & 64 & 44 & 28 & 36 & 40 & 49 & 25 & 35 & 35 & 64 \\
\bottomrule
\end{tabular}
\end{table}

@@

We observe that, aside from the Bross formula hdPS and Random Forest (RF) models, where the number of proxies was manually set to 100, and the kitchen sink (KS) model, which includes all proxies by design, the number of proxy variables selected by other models ranges between 49 and 69. As expected, the hybrid method combining Bross and LASSO hdPS selects exactly 49 proxy variables, all of which are selected by both methods, resulting in a common feature rate of 1.00. For other models, the common feature percentage is generally clustered around $74\%$, with XGBoost showing the highest common percentage at $79\%$, while the Genetic Algorithm (GA) displays the lowest common percentage at $69\%$.

>>>>

# Sample for structured abstract
abstract: |
  **Purpose**: We aim to evaluate various proxy selection methods within the context of high-dimensional propensity score (hdPS) analysis. This study aimed to systematically evaluate and compare the performance of traditional statistical methods and machine learning approaches within the hdPS framework, focusing on key metrics such as bias, standard error (SE), and coverage, under various exposure and outcome prevalence scenarios.
  **Methods:** We conducted a plasmode simulation study using data from the National Health and Nutrition Examination Survey (NHANES) cycles from 2013 to 2018. We compared methods including the kitchen sink model, Bross-based hdPS, Hybrid hdPS, LASSO, Elastic Net, Random Forest, XGBoost, and Genetic Algorithm (GA). The performance of each method was assessed based on bias, MSE, coverage probability, and SE estimation across three epidemiological scenarios: frequent exposure and outcome, rare exposure and frequent outcome, and frequent exposure and rare outcome.
  **Results:** XGBoost consistently demonstrated strong performance in terms of MSE and coverage, making it effective for scenarios prioritizing precision. However, it exhibited higher bias, particularly in rare exposure scenarios, suggesting it is less suited when minimizing bias is critical. In contrast, GA showed significant limitations, with consistently high bias and MSE, making it the least reliable method. Bross-based hdPS, and Hybrid hdPS methods provided a balanced approach, with low bias and moderate MSE, though coverage varied depending on the scenario. Rare outcome scenarios generally resulted in lower MSE and better precision, while rare exposure scenarios were associated with higher bias and MSE. Notably, traditional statistical approaches such as forward selection and backward elimination performed comparably to more sophisticated machine learning methods in terms of bias and coverage, suggesting that these simpler approaches may be viable alternatives due to their computational efficiency.
  **Conclusion:** The results highlight the importance of selecting hdPS methods based on the specific characteristics of the data, such as exposure and outcome prevalence. While advanced machine learning methods such as XGBoost can enhance precision, simpler methods such as forward selection or backward elimination may offer similar performance in terms of bias and coverage with fewer computational demands. Tailoring the choice of method to the epidemiological scenario is essential for optimizing the balance between bias reduction and precision.

@@

**High-dimensional Propensity Score (hdPS) Algorithm**: In epidemiology, proxy variables are commonly used as substitutes for confounders that are difficult or impossible to measure directly, such as socioeconomic status, lifestyle factors, or health behaviors [@vanderweele2019principles]. The hdPS is a pharmacoepidemiological method designed to reduce confounding bias in large healthcare databases [@schneeweiss2009high]. Unlike traditional propensity score models that rely on investigator-specified or manually selected covariates, hdPS automatically ranks a wide array of proxy variables from healthcare records—such as diagnosis codes, medications, and procedures—using the Bross formula [@wyss2018erratum; @bross1966spurious]. The Bross formula ranks these variables based on their marginal associations with both exposure and outcome. These selected proxy variables serve as surrogates for unmeasured or poorly measured confounders, helping reduce bias in treatment effect estimates. The hdPS algorithm further refines the selection by prioritizing variables based on their prevalence and potential for confounding, defined by their association with both exposure and outcome [@schneeweiss2009high].

**Multivariate Machine Learning Extensions**: Although the Bross formula performs well in certain contexts, it has limitations in capturing complex interactions, nonlinearities, and higher-order associations between variables, especially in high-dimensional settings where it does not account for the multivariate structure of other covariates [@karim2024high; @karim2018can]. To address these model-specification-related limitations, multivariate machine learning methods such as LASSO, Elastic Net, and Random Forests have been applied within the hdPS framework. These methods are better suited for high-dimensional data, where they can more effectively handle complex relationships and improve the selection of proxy variables, thus enhancing the precision of treatment effect estimates [@karim2018can; @schneeweiss2017variable; @franklin2015regularized]. Simulation studies and empirical research have shown that these machine learning-based methods, or hybrid approaches combining the Bross formula with machine learning, can reduce confounding more effectively and increase efficiency compared to the Bross formula alone in certain settings [@karim2018can; @schneeweiss2017variable; @franklin2015regularized].

**Assessing the Simulation Performance**: Other than bias, previous studies have primarily focused on Mean Squared Error (MSE) as a key metric for evaluating the performance of hdPS and its machine learning extensions [@franklin2015regularized; @pang2016effect; @wyss2018using; @karim2018can; @simon2023evaluating]. However, in high-dimensional settings with singly robust methods, such as hdPS and machine learning approaches such as LASSO, MSE may not always be the most reliable measure. MSE combines both bias and variance into a single metric, which makes interpretation challenging when variance estimation is unstable—a common issue in these methods. In contrast, coverage, which measures the proportion of confidence intervals that capture the true treatment effect, provides a more direct and meaningful assessment of a model's reliability. In realistic observational studies, where model misspecification is often inevitable, coverage—along with related metrics such as bias-eliminated coverage and relative error in standard error (SE; which compares model-based SE with empirical SE)—can reveal whether confidence intervals or SEs are too narrow (underestimating uncertainty) or too wide (overestimating uncertainty). This insight is crucial in determining whether the model delivers valid estimates despite misspecification. Even a model with poor MSE but good coverage may still be valuable, as it produces realistic confidence intervals. By shifting the focus to coverage, rather than relying solely on MSE, we can achieve a more comprehensive understanding of method performance, especially in cases where unstable variance estimation might distort conclusions drawn from MSE alone.

@@

**Plasmode simulation**: To rigorously assess the performance of the methods under consideration, we employed a plasmode simulation framework, which is particularly well-suited for reflecting real-world data structures and complexities [@franklin2014plasmode]. This approach was inspired by the analytic dataset derived from NHANES and involved resampling from the observed covariates and exposure information (i.e., obesity) without altering them. By mirroring key aspects of an actual epidemiological study, this simulation framework offers a substantial advantage over traditional Monte Carlo simulations, which often rely on hypothetical assumptions.

@@

**Performance Measures**: From this simulation, we derived several performance metrics to evaluate the effectiveness of the methods under consideration: (1) bias, (2) average model-based SE (the average of estimated SEs obtained from a model over repeated samples), (3) empirical SE (the standard deviation of estimated treatment effects across repeated samples), (4) MSE, (5) coverage probability of 95% confidence intervals, (6) bias-corrected coverage, and (7) Zip plot [@morris2019using; @white2023check].

@@

4. **Hybrid of hdPS and LASSO**: Instead of relying solely on LASSO for variable selection, a hybrid approach combines the Bross formula and LASSO. First, proxy variables are selected using the hdPS algorithm (e.g., the top 100), and then LASSO is applied to further refine the selection [@karim2018can; @franklin2015regularized].

@@

6. **Random Forest**: The Random Forest algorithm is an ensemble learning method that constructs multiple decision trees to perform classification [@breiman2001random]. It calculates the importance of each proxy variable based on the decrease in impurity or Gini importance, providing a ranking of the proxies. The top 100 variables from this ranking are manually selected for further modeling [@schneeweiss2017variable].

@@

9. **Genetic algorithm**: Genetic algorithm is an evolutionary algorithm inspired by the theory of natural selection [@holland1975adaptation]. It operates by evolving offspring from a population of the fittest individuals over several generations, evaluating and selecting the best combination of features or variables that maximize prediction accuracy.

@@

The results for each method under the different scenarios are summarized below. See Figures \ref{fig:bias-comparison} and \ref{fig:Coverage-comparison} for an overview of the performance in terms of bias and coverage, respectively. All simulation results can be reviewed interactively through a Shiny web application available at https://ehsanx.shinyapps.io/hdPS-Alternatives/, providing a convenient platform for exploring the performance of each method across various scenarios.

@@

1. *Bias*: Bross-based hdPS exhibited the smallest bias (-0.0001), and the kitchen sink model (0.0002) was the second. Genetic algorithm shows the highest bias (0.0287), indicating a substantial deviation from the true effect. Among the other methods, Hybrid hdPS (0.0016), and Elastic Net (0.0036) demonstrated low bias. XGBoost (0.0074) had slightly higher bias than Random Forest (0.0034).

2. *Coverage*: The coverage for most methods was high, with Hybrid hdPS, Forward Selection, Backward Elimination, LASSO, and Elastic Net achieving values around 98%, indicating well-calibrated confidence intervals. However, Genetic algorithm had noticeably lower coverage (83.8%), indicating that its confidence intervals might be too narrow or biased, potentially missing the true effect. After applying bias elimination (as there were substantial bias associated with this method), Genetic algorithm's coverage improved to 96%.

3. *MSE*: XGBoost achieved the lowest MSE (0.0006). Genetic algorithm maintained the highest MSE (0.0016), reflecting its higher bias and variability. The kitchen sink model (0.0009), Bross-based hdPS (0.0008), Hybrid hdPS (0.0008), and Elastic Net (0.0009) all had relatively similar and moderate MSE values.

4. *SE*: XGBoost exhibited the lowest Empirical SE (0.0229), indicating high precision in its estimates. The kitchen sink model had the highest Empirical SE (0.0305), suggesting greater variability. Other methods, including Genetic algorithm (0.0274), Hybrid hdPS (0.0278), and Bross-based hdPS (0.0287), showed moderate variability. LASSO (0.0299) and Elastic Net (0.0294) had slightly higher variability. The Model-based SE followed a similar pattern, with XGBoost (0.0268) showing the lowest variability and the kitchen sink model (0.0333) the highest, indicating less precision in its estimates. When comparing relative error in SE estimation, XGBoost performed the worst. See Appendix \S C for further details.

1. *Bias*: The kitchen sink model showed a relatively low bias (0.0025), while Genetic algorithm continued to exhibit the highest bias (0.0408), indicating a significant deviation from the true effect. XGBoost had a bias of 0.0259, which, while still higher than other methods, but was lower than Genetic algorithm. Other methods, such as Bross-based hdPS (0.0035), Hybrid hdPS (0.0049), and Elastic Net (0.0053), demonstrated moderate bias. Random Forest (0.0127) and Forward Selection (0.0108) had slightly higher bias but remained within an acceptable range.

2. *Coverage*: Coverage levels remained high for most methods, with XGBoost achieving the highest coverage at 96.2%, indicating well-calibrated confidence intervals despite its higher bias. The Genetic algorithm method had lower coverage (92.2%), suggesting that its confidence intervals might be narrower, potentially missing the true effect. Other methods such as RF, Forward Selection, Backward Elimination, and Hybrid hdPS maintained coverage values around 94-95%, suggesting adequate interval calibration. Bias-eliminated coverage for Genetic algorithm improved to 94.2%, but it was still slightly lower than other methods (e.g., forward selection).

3. *MSE*: Forward selection demonstrated the lowest MSE (0.0030), and then Hybrid hdPS and XGBoost. The Genetic algorithm method had a higher MSE (0.0043), reflecting its substantial bias and variability. The kitchen sink model also had an MSE of 0.0043, similar to Genetic algorithm, while other methods such as Bross-based hdPS (0.0035), RF (0.0039), and Elastic Net (0.0036) exhibited moderate MSE values, indicating reasonable accuracy.

4. *SE*: The lowest Empirical SE was observed with XGBoost (0.0507) and Genetic algorithm (0.0510), reflecting high precision despite their higher bias. The kitchen sink model had the highest Empirical SE (0.0656), indicating greater variability. Hybrid hdPS (0.0564), Bross-based hdPS (0.0595), and RF (0.0609) showed moderate variability. Forward Selection (0.0537) and Backward Elimination (0.0576) had lower variability compared to the kitchen sink model. In terms of Model-based SE, XGBoost (0.0531) and Genetic algorithm (0.0533) continued to show low variability, while the kitchen sink model had the highest Model-based SE (0.0623), indicating less precision in its estimates. When comparing relative error in SE estimation, XGBoost and kitchen sink model performed the worst (in other direction).

1. *Bias*: In this scenario, the kitchen sink model exhibited a moderate negative bias (-0.0093), similar to the Bross-based hdPS method (-0.0088). Genetic algorithm showed a significantly higher bias (0.0362), indicating a substantial deviation from the true effect. Among other methods, XGBoost demonstrated the lowest bias (-0.0061), while methods such as Hybrid hdPS (-0.0082), Forward Selection (-0.0070), and Backward Elimination (-0.0070) had slightly higher but still moderate biases. Elastic Net and LASSO both had biases of -0.0079, reflecting slightly larger deviations compared to XGBoost but still within acceptable limits.

2. *Coverage*: Most methods achieved good coverage, with XGBoost, RF, and Forward Selection each achieving a coverage rate of 95.4%, indicating well-calibrated confidence intervals. The Genetic algorithm method, however, had slightly lower coverage (91.8%), indicating that its confidence intervals might be narrower, potentially excluding the true effect. Bross-based hdPS and the kitchen sink model had slightly lower coverage values of 93.8% and 93.4%, respectively. After accounting for bias, the bias-eliminated coverage for most methods, except Genetic algorithm, remained high, with values ranging from 98.4% to 99.0%, indicating that most methods effectively adjusted for bias in their coverage estimates. Genetic algorithm's bias-eliminated coverage was lower at 93.4%, reflecting its higher inherent bias.

3. *MSE*: XGBoost exhibited the lowest MSE (0.0003). Genetic algorithm had the highest MSE (0.0040), reflecting its substantial bias and variability. The kitchen sink model (0.0005), Bross-based hdPS (0.0005), and other methods such as Hybrid hdPS (0.0004) and Elastic Net (0.0005) all had relatively similar MSE values, indicating moderate accuracy.

4. *SE*: The lowest Empirical SE was observed with XGBoost (0.0152), reflecting high precision in its estimates. The Genetic algorithm method exhibited the highest Empirical SE (0.0523), indicating greater variability and less precision. Methods such as Hybrid hdPS (0.0184), Forward Selection (0.0187), and Elastic Net (0.0203) showed moderate variability, while Bross-based hdPS (0.0206) and the kitchen sink model (0.0212) had slightly higher variability. In terms of Model-based SE, XGBoost (0.0179) again showed the lowest variability, consistent with its low Empirical SE, indicating that it provided the most stable estimates. The kitchen sink model had a slightly higher Model-based SE (0.0219), indicating less precision in its estimates. When comparing relative error in SE estimation, XGBoost performed the worst.

@@

# Combine RD and OR data into a single data frame
data_combined <- data.frame(
  Method = rep(c("Kitchen Sink", "Bross", "Hybrid", "LASSO", "Elastic Net", "Random Forest", "XGBoost", "Fprward Selection", "Backward Elimination", "Genetic algorithm"), 2),
  # These are approximate values
  Value = round(c(0.075, 0.077, 0.079, 0.078, 0.075, 0.085, 0.082, 0.078, 0.078, 0.076, 
            1.502, 1.522, 1.543, 1.531, 1.504, 1.588, 1.564, 1.533, 1.533, 1.515),2),
  LowerCI = round(c(0.056, 0.058, 0.06, 0.059, 0.056, 0.066, 0.063, 0.059, 0.059, 0.057,
              1.395, 1.413, 1.432, 1.421, 1.396, 1.474, 1.452, 1.423, 1.423, 1.406),2),
  UpperCI = round(c(0.095, 0.096, 0.099, 0.098, 0.094, 0.104, 0.101, 0.097, 0.097, 0.096,
              1.618, 1.64, 1.663, 1.65, 1.62, 1.712, 1.685, 1.653, 1.653, 1.632),2),
  Measure = rep(c("RD", "OR"), each = 10),
  Variables = rep(c(142, 100, 49, 60, 69, 100, 48, 59, 59, 64), 2)
)

@@

\begin{table}[htbp]
\centering
\caption{Comparison of variable overlap of selected proxies across different methods used to evaluate the association between obesity and diabetes from the National Health and Nutrition Examination Survey (NHANES) for the years 2013-2018.}
\label{tab:method-comparison-updated}
\begin{tabular}{lcccccccccc}
\toprule
 & \textbf{KS} & \textbf{Bross} & \textbf{Hybrid} & \textbf{LASSO} & \textbf{EN} & \textbf{RF} & \textbf{XGB} & \textbf{FS} & \textbf{BE} & \textbf{GA} \\
\midrule
\textbf{Kitchen sink (KS)} & 142 & & & & & & & & & \\
\textbf{Bross formula} & 100 & 100 & & & & & & & & \\
\textbf{Hybrid (Bross and LASSO)} & 49 & 49 & 49 & & & & & & & \\
\textbf{LASSO} & 60 & 47 & 47 & 60 & & & & & & \\
\textbf{Elastic Net (EN)} & 69 & 54 & 48 & 60 & 69 & & & & & \\
\textbf{Random Forest (RF)} & 100 & 71 & 38 & 46 & 52 & 100 & & & & \\  
\textbf{XGBoost (XGB)} & 48 & 38 & 24 & 28 & 30 & 37 & 48 & & & \\
\textbf{Forward selection (FS)} & 59 & 45 & 41 & 51 & 54 & 45 & 25 & 59 & & \\
\textbf{Backward elimination (BE)} & 59 & 45 & 41 & 51 & 54 & 45 & 25 & 59 & 59 & \\
\textbf{Genetic algorithm (GA)} & 64 & 44 & 28 & 36 & 40 & 49 & 25 & 35 & 35 & 64 \\
\bottomrule
\end{tabular}
\end{table}

@@

We observe that, aside from the Bross formula hdPS and Random Forest models, where the number of proxies was manually set to 100, and the kitchen sink (KS) model, which includes all proxies by design, the number of proxy variables selected by other models ranges between 49 and 69. As expected, the hybrid method combining Bross and LASSO hdPS selects exactly 49 proxy variables, all of which are selected by both methods, resulting in a common feature rate of 1.00. For other models, the common feature percentage is generally clustered around $74\%$, with XGBoost showing the highest common percentage at $79\%$, while the Genetic Algorithm (GA) displays the lowest common percentage at $69\%$.

\begin{table}[htbp]
\centering
\caption{Comparison of the count and percentage of proxy variables selected by each methods in common with that by the Bross formula-based high-dimensional propensity score to evaluate the association between obesity and diabetes from the National Health and Nutrition Examination Survey (NHANES) for the years 2013-2018.}
\label{tab:proxy-common-pct-bross}
\begin{tabular}{lccc}
\toprule
\textbf{Method} & \textbf{Total Count} & \textbf{Common Count} & \textbf{Rate in Common} \\ 
\midrule
\textbf{Kitchen sink} & 142 & - & - \\
\textbf{Bross formula} & 100 & - & - \\
\textbf{Hybrid (Bross and LASSO)} & 49 & 49 & 1.00 \\
\textbf{LASSO} & 60 & 47 & 0.78 \\
\textbf{Elastic Net} & 69 & 54 & 0.78 \\
\textbf{Random Forest} & 100 & 71 & 0.71 \\  
\textbf{XGBoost} & 48 & 38 & 0.79 \\
\textbf{Forward selection} & 59 & 45 & 0.76 \\
\textbf{Backward elimination} & 59 & 45 & 0.76 \\
\textbf{Genetic algorithm} & 64 & 44 & 0.69 \\
\bottomrule
\end{tabular}
\end{table}
